Named Entity Recognition using Structured Prediction

### 2012-11-03 doc/report/paper.tex
Trees and so on

\documentclass[11pt]{article}
\usepackage{acl2011}
\usepackage[utf8]{inputenc}
\usepackage{linguex}
\usepackage{latexsym}  
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{amsmath}
% \usepackage{xcolor}
% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
% \definecolor{medium-blue}{rgb}{0,0,0.5}
% \hypersetup{
%     colorlinks, linkcolor={dark-red},
%     citecolor={dark-blue}, urlcolor={medium-blue}}
\title{Named Entity Recognition using Structured Prediction}

@@

@@

In considering named entities, it is important to distinguish between mention of entities that are non - specific as 
time expressions, such as \emph{In June}, referring to any possible year or \emph{the prof} as a person name , which itself is not a specific entity, but a deictic reference
, which may point back to the mention of a real \emph{NE}, as in the following coreference chain: \\

\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$

@@

Prominent algorithms in NER include maximum entropy
% Overview algorithms /methods used 
% machine learning/neural networks/rule-based 
% maximum entropy MM
%

\subsection*{CoNLL data set \& baseline}
The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{TjongKimSang:2003:ICS:1119176.1119195} and the \emph{CoNLL Shared Task 2003} \cite{tksintro2002conll}
for Spanish \& Dutch and English \& German respectively.

% A baseline rate was computed for the English and the
% German test sets. It was produced by a system which
% only identified entities which had a unique class in
% the training data. If a phrase was part of more than
% one entity, the system would select the longest one.
% All systems that participated in the shared task have
% outperformed the baseline system.
% For all the Fβ=1 rates we have estimated sig-
% nificance boundaries by using bootstrap resampling
% (Noreen, 1989). From each output file of a system,
% 250 random samples of sentences have been chosen
% and the distribution of the Fβ=1 rates in these sam-
% ples is assumed to be the distribution of the perfor-
% mance of the system. We assume that performance
% A is significantly different from performance B if A
% is not within the center 90% of the distribution of B.

@@

\section{Structured Prediction}
Structured Prediction \cite{strlearn} 
is a supervised-learning approach and sets itself apart from Non-structured Prediction through the form of its output. 
Prediction maps an input x to an output y: \\
$ x \rightarrow y $.

Non-structured output is atomic, thus it is binary prediction for a two-class problem and may corresponds to more than one of more than 2 possible labels for a 
multiclass problem. 
The output of Structured Prediction is structured and gives back a sequence/tree. 
Thus, the output could be the complete sentence with its corresponding labels.

\ex. \emph{The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903.} \label{PredEx1a}

would trigger the labels: 
\ex. x The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903. \label{PredEx1b} % include labels 
     %y 
Furthermore, Structured Prediction is different from similar approaches such as HMM through the combination of features and the use of label interactions.

\section{Our structured Perceptron}
Our NER structured perceptron for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
respectively.

\paragraph*{Learning}
Labels of the whole sentence (0) for zero entity
\paragraph*{Structure}
Structured Perceptron with Averaging

\paragraph*{Decoding}
Viterbi algorithm (Markov assumption, only 1 prev.
label)

@@

\begin{table}[h!]
\begin{tabular}{l|l}
\hline
\bf Feature & \bf Example \\
\hline
Token &  \\
Suffix& Amster\textbf{dam}\\
Prefix&  \textbf{San} Sebastian\\
Captitalized& \textbf{B}enetton\\
Number Pattern & \\
UPPERCASE &  \textbf{BENETTON}\\
POS-tag &  Benetton \textbf{NNP}   \\
Lemma &   \\
\end{tabular}
\caption{Node Features}
\label{table:node}
\end{table}

\paragraph*{Label Interaction Features}
These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
%current token and last label for prepositions or possessive ’s

@@

\begin{table}[h!]
\begin{tabular}{l |l| l| l| l| l| l}

\bf Language & \multicolumn{3}{c|}{ \bf testA}&\multicolumn{3}{c}{ \bf testB}\\\hline 
             & Precision & Recall & $F_1$-measure & Precision & Recall & $F_1$-measure \\ \hline
Spanish &       &          &     &          &               & \\
Dutch  &         &          &     &          &               &   \\
English &        &          &     &          &               &       \\
German &      &          &       &          &             & \\
\end{tabular}
\caption{NER Structured Prediction Results }
\label{table:Results}
\end{table}

>>>>

\documentclass[11pt]{article}
\usepackage{acl2011}
\usepackage[utf8]{inputenc}
\usepackage{linguex}
\usepackage{latexsym}  
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{qtree}

% \usepackage{xcolor}
% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
% \definecolor{medium-blue}{rgb}{0,0,0.5}
% \hypersetup{
%     colorlinks, linkcolor={dark-red},
%     citecolor={dark-blue}, urlcolor={medium-blue}}
\title{Named Entity Recognition using Structured Prediction}

@@

\DeclareMathOperator*{\argmax}{arg\,max}

@@

In considering named entities, it is important to distinguish between mention of entities that are non-specific as 
time expressions, such as \emph{In June}, referring to any possible year or \emph{the prof} as a person name, which itself is not a specific entity, 
but a deictic reference, which may point back to the mention of a real \emph{NE}, as in the following coreference chain: \\

\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \\

Named Entities can consist of more than one syntactic type. The usual overall type for a named entity is noun phrase, which can be both simple and complex. 
An example for a simple heterogenous entity is displayed in syntactic tree represenation in ~\ref{tree1}.
 \Tree
    [.NP [.N' John ]  ] \label{tree1}

Most languages allow for a recursive definition of, for instance, a noun phrase that in theory can grow infinitely big. 
An \textbf{NP} can consist of a determiner, such as \emph{the} and a common noun or simply a propernoun. \textbf{PP} stands
for prepositional phrase and consists of a preposition and a noun phrase. 
The syntactic rules for PP-attachment are displayed in \ref{PP}.
\begin{figure}
$ NP \rightarrow NP PP$ \\
$ PP \rightarrow P NP$ 
\caption{Recursive Noun Phrase}
\label{PP}
\end{figure}   
These rules allow for constructions such as in sentence~\ref{PPrec}.

\ex. The princess of Scotland is in the castle in England with the brown bricks on its roof... \label{PPrec}

We distinguish between PP-complements and PP-modifiers, where modifiers simply add to the description of the noun phrase, maybe
to set it apart and complements are more inherent to the meaning of the entity. 
The PP-attachment \emph{of Scotland} is semantically defining for the noun \emph{princess}, whereas \emph{with the brown bricks on its roof} 
just adds to the description of \emph{the castle in England}.
Regarding named entities, one wants to differentiate between a PP-attachment that defines the entity an one that is irrelevant.

\begin{enumerate}
\item \parbox[t]{2.4in}{ (a)~\Tree 
   [.NP [ [.DT the ] [.N\1 [.N\1 Princess ] [.PP [.P of ] [.NP England ] ] ] ] ]}
\parbox[t]{2in}{ (b)~\Tree 
   [.NP  [.DT the ] [.N\1 [.N\1 Princess ] ] [.PP [.P from ] [.NP England ] ]  ] ]  }
\end{enumerate}

@@

Prominent algorithms in NER include maximum entropy models,

neural network

\subsection*{CoNLL Data Set \& Baseline Performance}
The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
% A baseline rate was computed for the English and the
% German test sets. It was produced by a system which
% only identified entities which had a unique class in
% the training data. If a phrase was part of more than
% one entity, the system would select the longest one.
% All systems that participated in the shared task have
% outperformed the baseline system.
% For all the Fβ=1 rates we have estimated sig-
% nificance boundaries by using bootstrap resampling
% (Noreen, 1989). From each output file of a system,
% 250 random samples of sentences have been chosen
% and the distribution of the Fβ=1 rates in these sam-
% ples is assumed to be the distribution of the perfor-
% mance of the system. We assume that performance
% A is significantly different from performance B if A
% is not within the center 90% of the distribution of B.

@@

\section{Structured Prediction}
Structured Prediction \cite{strlearn} is a supervised-learning approach that maps an input \textbf{x} to an output \textbf{y}: \\

$ x \rightarrow y $. \\

Structured and  Non-structured prediction differ in the form of their output. 
Non-structured output is atomic; it is binary prediction for a two-class problem and may corresponds to more than one of more than two possible labels for a 
multiclass problem. 
Structured Prediction gives back the prediction for the whole sequence. 
For the sentence in \ref{PredEx1a}, it would return the corresponding labels combination in example~\ref{PredEx1b}.

\ex. \emph{The motor company 'Ford' was founded and incorporated  by Henry Ford.} \label{PredEx1a}
%The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903.

\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} 0  0  ORG  0  0 0 0 0 PER PER.  \label{PredEx1b}

Also, Structured Prediction is different from similar approaches through its combination of features and label interactions instead
of concentrating mainly on label interactions like HMM or a rich set of features like local classifiers.

\section{Our Implementation}
% problem external data / use Dbpedia - gazetteer
Our NER system for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
respectively.

\paragraph*{Learning}
Labels of the whole sentence (0) for zero entity

\paragraph*{Structure}
Structured Perceptron with Averaging

\paragraph*{Decoding}
In this system we are using viterbi decoding, thus we are looking for the prediction $\hat{y}$, which is the sequence with the highest overall score:

\[
\hat{y} = \argmax_{\mathbf{y} \in \mathcal{Y}^{n}} \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})
\]

This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ using the Viterbi algorithm. 
As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
The score of the best sequence ending in $a$ is:

\[
\delta_i(a) = \max_{\mathbf{y} \in \mathcal{Y}^{n}, y_n = a} \sum_{j=1}^{n}{\mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, j, y_{j-1}, y_{j})}
\]

\noindent And $\delta_i(a)$ can be calculated recursively:

\begin{align*}
\delta_1(a) &= \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, 1, \emptyset, a) \\
\delta_i(a) &= \max_{b \in \mathcal{Y}} \delta_{i-1}(b) + \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, b, a) \\
\end{align*}

@@

\begin{table}[h!]
\begin{tabular}{l|l}
\hline
\bf Feature & \bf Example \\
\hline
Token &  \textbf{``Amsterdam''}\\
Suffix& Amster\textbf{dam}\\
Prefix&  \textbf{San} Sebastian\\
Captitalized& \textbf{B}enetton\\
Number Pattern & \\
UPPERCASE &  \textbf{BENETTON}\\
POS-tag &  Benetton \textbf{NNP}   \\
Lemma &  produced $Rightarrow$ produce \\
\end{tabular}
\caption{Node Features}
\label{table:node}
\end{table}

\paragraph*{Label Interaction Features}
These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1}is more likely 
than\emph{PER} and \emph{LOC} as presented in \ref{seq2}.
.
%current token and last label for prepositions or possessive ’s

\ex. Jack London went to New York. \\ PER   PER  0  0  LOC   LOC. \label{seq1}

\ex. Jack London went to New York. \\ PER   LOC  0  0  LOC   LOC. \label{seq2}

@@

\begin{table}[h!]
\scriptsize
\begin{tabular}{l |l| l| l| l| l| l}

\bf Language & \multicolumn{3}{c|}{ \bf testA}&\multicolumn{3}{c}{ \bf testB}\\\hline 
             & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ \hline
Spanish &       &          &     &          &               & \\
Dutch  &         &          &     &          &               &   \\
English &        &          &     &          &               &       \\
German &      &          &       &          &             & \\
\end{tabular}
\caption{NER Structured Prediction Results }
\label{table:Results}
\end{table}
